%!TEX root = ../manyExamples.tex
\subsection{\texttt{heatXchanger}: Local analysis of heat exchanger} 
\label{heatXchanger}


\cite{Roberts2013a} provides novel theoretical support for the method of multiple scales in spatio-temporal systems, and then extends this important method.
Perhaps the simplest example is the heat exchanger: the non-autonomous slow manifold analysis that is at the heart of the novel methodology is determined here.
An internal technical report\footnote{AJR (2012), \emph{Derive boundary conditions for heat exchanger modelling using near boundary dynamics}, \texttt{dbefhem.pdf}} considered the scenario where hot fluid enters one pipe from the right having temperature field~$a(x,t)$, and cold fluid enters the other pipe from the left with temperature field~$b(x,t)$.  
Non-dimensional governing \pde{}s are
\begin{equation*}
\D ta=+\D xa+\rat12(b-a),\quad
\D tb=-\D xb+\rat12(a-b).
\label{eq:hepde}
\end{equation*}

Transform to mean and difference fields:
\begin{equation*}
c(x,t):=\rat12(a+b),\quad d(x,t):=\rat12(a-b),
\quad\text{i.e., }
a=c+d,\quad b=c-d.
\label{eq:meandiff}
\end{equation*}
The mean and difference of the \pde{}s gives the equivalent \pde\ system
\begin{equation*}
\D tc=\D xd,\quad
\D td=-d+\D xc.
\label{eq:hecan}
\end{equation*}
In this form we readily see that the difference field~$d$ tends to decay exponentially quickly, but that interaction between gradients of the mean and difference fields generates other effects.

The approach is to expand the fields in their local spatial structure based around a station $x=X$.
Expand advection-exchange in a heat exchanger in powers of $(x-X)^n/n!$.
\begin{align*}
c(x,t)&=c_0(X,t)+c_1(X,t)(x-X) +c_2(X,t)\rat12(x-X)^2
\nonumber\\&\quad{}
+c_3(X,t)\rat16(x-X)^3
 +c_4(X,x,t)\rat1{24}(x-X)^4,
%\label{eq:ctrt}
\\
d(x,t)&=d_0(X,t)+d_1(X,t)(x-X) +d_2(X,t)\rat12(x-X)^2
\nonumber\\&\quad{}
+d_3(X,t)\rat16(x-X)^3 
+d_4(X,x,t)\rat1{24}(x-X)^4,
%\label{eq:dtrt}
\end{align*}
With Taylor Remainder Theorem closing the problem in terms of unknown functions which here are represented by the non-autonomous forcing~$w_i$.
Variables \(\verb|y(j)|=d_{j-1}\) and \(\verb|x(j)|=c_{j-1}\).  
Also \(\verb|w(1)|=d_{4X}\eta_x\) and \(\verb|w(2)|=c_{4X}\xi_x\) and evaluate at intensity $\sigma=5$\,.

Start by loading the procedure.
\begin{reduce}
in_tex "../stoNormForm.tex"$
\end{reduce}
Execute the construction of a normal form for the system.
\begin{reduce}
stonormalform(
    {y2,y3,y4,y5,w(1)},
    {-y1+x2
    ,-y2+x3
    ,-y3+x4
    ,-y4+x5
    ,-y5+w(2)},
    { },
    99 );
%end; % optional finish here
\end{reduce}
Here, since the results are exact, we can notionally carry out analysis to high-order, here coded~\(99\)th order.  Alternatively, as in the next section, one could divide by~\verb|small| all the~\verb|y(j)| terms in the \verb|x(i)| equations, and all the~\verb|x(i)| terms in the~\verb|y(j)| equations to analyse the original system without the algorithm's artifice of~\verb|small|/\eps.

\paragraph{Specified dynamical system}
The above embeds the \ode{}s as the following. 
\begin{align*}&
\dot x_{1}=\varepsilon  y_{2}
\\& 
\dot x_{2}=\varepsilon  y_{3}
\\& 
\dot x_{3}=\varepsilon  y_{4}
\\& 
\dot x_{4}=\varepsilon  y_{5}
\\& 
\dot x_{5}=\sigma  w_{1}
\\& 
\dot y_{1}=\varepsilon  x_{2}-y_{1}
\\& 
\dot y_{2}=\varepsilon  x_{3}-y_{2}
\\& 
\dot y_{3}=\varepsilon  x_{4}-y_{3}
\\& 
\dot y_{4}=\varepsilon  x_{5}-y_{4}
\\& 
\dot y_{5}=\sigma  w_{2}-y_{5}
\end{align*}

\paragraph{Time dependent coordinate transform}
\begin{align*}&
y_{1}=\ParMath{ \sigma  \eps^{4} \big(\ou\big(\ou\big(\ou\big(w_{2},tt,-1\big),tt,
-1\big),tt,-1\big)+2 \ou\big(\ou\big(w_{2},tt,-1\big),tt,-1\big)+3 \ou
\big(w_{2},tt,-1\big)\big)-\eps^{3} X_{4}+\eps X_{2}+Y_{1}
}\\& 
y_{2}=\sigma  \eps^{3} \big(\ou\big(\ou\big(w_{1},tt,-1\big),tt,-1\big)+
2 \ou\big(w_{1},tt,-1\big)\big)-\eps^{3} X_{5}+\eps X_{3}+Y_{2}
\\& 
y_{3}=\sigma  \eps^{2} \big(-\ou\big(\ou\big(w_{2},tt,-1\big),tt,-1\big)
-\ou\big(w_{2},tt,-1\big)\big)+\eps X_{4}+Y_{3}
\\& 
y_{4}=-\sigma  \eps \ou\big(w_{1},tt,-1\big)+\eps X_{5}+Y_{4}
\\& 
y_{5}=\sigma  \ou\big(w_{2},tt,-1\big)+Y_{5}
\\& 
x_{1}=\sigma  \eps^{4} \big(-\ou\big(\ou\big(w_{1},tt,-1\big),tt,-1\big)
-3 \ou\big(w_{1},tt,-1\big)\big)+\eps^{3} Y_{4}-\eps Y_{2}+X_{1}
\\& 
x_{2}=\sigma  \eps^{3} \big(\ou\big(\ou\big(w_{2},tt,-1\big),tt,-1\big)+
2 \ou\big(w_{2},tt,-1\big)\big)+\eps^{3} Y_{5}-\eps Y_{3}+X_{2}
\\& 
x_{3}=\sigma  \eps^{2} \ou\big(w_{1},tt,-1\big)-\eps Y_{4}+X_{3}
\\& 
x_{4}=-\sigma  \eps \ou\big(w_{2},tt,-1\big)-\eps Y_{5}+X_{4}
\\& 
x_{5}=X_{5}
\end{align*}


\paragraph{Result normal form DEs}
\begin{align*}&
\dot Y_{1}=\eps^{4} Y_{5}-\eps^{2} Y_{3}-Y_{1}
\\& 
\dot Y_{2}=-\eps^{2} Y_{4}-Y_{2}
\\& 
\dot Y_{3}=-\eps^{2} Y_{5}-Y_{3}
\\& 
\dot Y_{4}=-Y_{4}
\\& 
\dot Y_{5}=-Y_{5}
\\& 
\dot X_{1}=3 \sigma  \eps^{4} w_{1}-\eps^{4} X_{5}+\eps^{2} X_{3}
\\& 
\dot X_{2}=-2 \sigma  \eps^{3} w_{2}+\eps^{2} X_{4}
\\& 
\dot X_{3}=-\sigma  \eps^{2} w_{1}+\eps^{2} X_{5}
\\& 
\dot X_{4}=\sigma  \eps w_{2}
\\& 
\dot X_{5}=\sigma  w_{1}
\end{align*}
Clearly, the emergent slow manifold is \(\Yv=0\).
Then the slow evolution of~\Xv\ leads to the emergent dynamics being described by 
\begin{equation*}
\D tc\approx\DD xc-\Dn x4c+3\sigma w_1\,.
\end{equation*}



\subsubsection{Near the boundary}
This is for the case of boundary conditions $c+pd=\text{cd}_0(t)$ at $x=0$ for some parameter~$p$.  
Computer algebra finds boundary conditions on the fields that reduce the dynamics near the boundary to the following with \(\verb|x(1)|=c_1\), \(\verb|x(2)|=c_3\), \(\verb|y(1)|=d_0\), \(\verb|y(2)|=d_2\) and \(\verb|w(1)|=d_{3X}\eta_x\) with \(\sigma=4\).
Curiously, there is no dependence upon parameter~$p$ in these dynamics.

The procedure \verb|stonormalform| is already loaded.
Write a message saying we are now analysing the next system.
\begin{reduce}
write "**** Near the boundary ****";
\end{reduce}
Execute the construction of a normal form for this system
\begin{reduce}
stonormalform(
    {y(2)/small,w(1)},
    {-y(1)+x(1)/small,-y(2)+x(2)/small},
    { },
    99 );
%end; % optional finish here
\end{reduce}
Again, the following results are exact.

\paragraph{Specified dynamical system}
\begin{align*}&
\dot x_{1}=y_{2}
\\&
\dot x_{2}=\sigma  w_{1}
\\&
\dot y_{1}=x_{1}-y_{1}
\\&
\dot y_{2}=x_{2}-y_{2}
\end{align*}

\paragraph{Time dependent coordinate transform}
\begin{align*}&
y_{1}=\sigma  \big(\ou\big(\ou\big(w_{1},tt,-1\big),tt,-1\big)+2 \ou
\big(w_{1},tt,-1\big)\big)-X_{2}+X_{1}+Y_{1}
\\&
y_{2}=-\sigma  \ou\big(w_{1},tt,-1\big)+X_{2}+Y_{2}
\\&
x_{1}=\sigma  \ou\big(w_{1},tt,-1\big)+X_{1}-Y_{2}
\\&
x_{2}=X_{2}
\end{align*}


\paragraph{Result normal form DEs}
\begin{align*}&
\dot Y_{1}=-Y_{2}-Y_{1}
\\&
\dot Y_{2}=-Y_{2}
\\&
\dot X_{1}=-\sigma  w_{1}+X_{2}
\\&
\dot X_{2}=\sigma  w_{1}
\end{align*}




\subsubsection{Heat exchanger with quadratic reaction}
\label{sec:heqr}

Expand advection-reaction-exchange in a heat exchanger in powers of $(x-X)^n/n!$.  
The reaction is some quadratic that should generate Burgers' equation model.
With Taylor Remainder Theorem closing the problem in terms of unknown functions which here are represented by the non-autonomous forcing~$w_i$.
Note that \(\verb|y(j)|=d_{j-1}\) and \(\verb|x(j)|=c_{j-1}\).  
Also \(\verb|w(1)|=3d_{2x}\) and \(\verb|w(2)|=3c_{2x}\) and evaluate at intensity $\sigma=1$\,.

The procedure \verb|stonormalform| is already loaded.
Write a message saying we are now analysing the next system.
\begin{reduce}
write "**** with quadratic reaction ****";
\end{reduce}
Execute the construction of a normal form for this system, but this time just choose the slow centre manifold.
\begin{reduce}
stonormalform(
    { y(2)-x(1)*y(1),
      y(3)-x(1)*y(2)-x(2)*y(1),
      small*w(1)-x(1)*y(3)-2*x(2)*y(2)-x(3)*y(1) },
    { -y(1)+x(2)-(x(1)^2+y(1)^2)/2,
      -y(2)+x(3)-x(1)*x(2)-y(1)*y(2),
      -y(3)+small*w(2)-x(2)^2-x(1)*x(3)-y(2)^2-y(1)*y(3) },
    { },
    3 , cman);
end;
\end{reduce}

We could divide the off-diagonal linear terms by~\verb|small| (and remove the multiplication of forcing~\verb|w|), and the algorithm still converges, albeit in more iterations.  
The resulting asymptotic expressions then do not assume that \(x\)~derivatives are `successively smaller'.  
\footnote{See result in file \texttt{sdeheqr.pdf}}
The following uses the default scaling which corresponds to successively smaller \(x\)-derivatives provided I also multiply the forcing by~\verb|small|.

\paragraph{Specified dynamical system}
\begin{align*}&
\dot x_{1}=\eps \big(-x_{1} y_{1}+y_{2}\big)
\\&
\dot x_{2}=\eps \big(-x_{2} y_{1}-x_{1} y_{2}+y_{3}\big)
\\&
\dot x_{3}=\sigma  \eps w_{1}+\eps \big(-x_{3} y_{1}-2 x_{2} y_{2}-x_{1}
 y_{3}\big)
\\&
\dot y_{1}=\eps \big(x_{2}-1/2 x_{1}^{2}-1/2 y_{1}^{2}\big)-y_{1}
\\&
\dot y_{2}=\eps \big(x_{3}-x_{2} x_{1}-y_{2} y_{1}\big)-y_{2}
\\&
\dot y_{3}=\sigma  \eps w_{2}+\eps \big(-x_{3} x_{1}-x_{2}^{2}-y_{3} y_{
1}-y_{2}^{2}\big)-y_{3}
\end{align*}


\paragraph{Time dependent coordinate transform}
\begin{align*}&
y_{1}=\eps \big(X_{2}-1/2 X_{1}^{2}
%+1/2 Y_{1}^{2}
\big)%+Y_{1}
\\&
y_{2}=\eps \big(X_{3}-X_{2} X_{1}
%+Y_{2} Y_{1}
\big)%+Y_{2}
\\&
y_{3}=\sigma  \eps \ou\big(w_{2},tt,-1\big)+\eps \big(-X_{3} X_{1}-X_{2}
^{2}
%+Y_{3} Y_{1}+Y_{2}^{2}
\big)%+Y_{3}
\\&
x_{1}=%\eps \big(X_{1} Y_{1}-Y_{2}\big)+
X_{1}
\\&
x_{2}=%\eps \big(X_{2} Y_{1}+X_{1} Y_{2}-Y_{3}\big)+
X_{2}
\\&
x_{3}=%\eps \big(X_{3} Y_{1}+2 X_{2} Y_{2}+X_{1} Y_{3}\big)+
X_{3}
\end{align*}

\paragraph{Result normal form DEs}
\begin{align*}&
%\dot Y_{1}=\eps^{2} \big(-1/2 X_{1}^{2} Y_{1}+2 X_{1} Y_{2}-Y_{3}\big)-Y
%_{1}
%\\&
%\dot Y_{2}=\eps^{2} \big(-X_{2} X_{1} Y_{1}+2 X_{2} Y_{2}-1/2 X_{1}^{2} 
%Y_{2}+2 X_{1} Y_{3}\big)-Y_{2}
%\\&
%\dot Y_{3}=\ParMath{ -\sigma  \eps^{2} w_{2} Y_{1}+\eps^{2} \big(-X_{3} X_{1} Y_{1
%}-X_{3} Y_{2}-X_{2}^{2} Y_{1}-2 X_{2} X_{1} Y_{2}+X_{2} Y_{3}-1/2 X_{1}
%^{2} Y_{3}\big)-Y_{3}
%}\\&
\dot X_{1}=\eps^{2} \big(X_{3}-2 X_{2} X_{1}+1/2 X_{1}^{3}\big)
\\&
\dot X_{2}=\sigma  \eps^{2} w_{2}+\eps^{2} \big(-2 X_{3} X_{1}-2 X_{2}^{
2}+3/2 X_{2} X_{1}^{2}\big)
\\&
\dot X_{3}=-\sigma  \eps^{2} w_{2} X_{1}+\sigma  \eps w_{1}+\eps^{2} 
\big(-3 X_{3} X_{2}+3/2 X_{3} X_{1}^{2}+3 X_{2}^{2} X_{1}\big)
\end{align*}
Hmmm, looks like this generates the slowly varying model that
\begin{equation*}
\D tC\approx \DD xC-2C\D xC+\rat12C^3.
\end{equation*}
Interestingly there is an extra factor of two in the nonlinear advection, and a net cubic reaction.


