%!TEX root = ../manyExamples.tex
\subsection{\texttt{othersFail}: Other methodologies fail} 
\label{othersFail}

Consider, for small bifurcation parameter~$\epsilon$, the system
\begin{align*}
\text{slow mode }& \dot x=\epsilon x+x^3-xy +\sigma xyw\,,\\
\text{fast mode }& \dot y= -y+x^2+y^2+\sigma yw\,.
\end{align*}
Deterministically, there is a bifurcation to two equilibria for small $\epsilon>0$\,.
The noise~$w$ affects this bifurcation somehow.

Why is this tricky? \emph{Cross-sectional averaging} is simply projection onto the slow space $y=0$ which predicts instability of subcritical bifurcation $\dot x=\epsilon x+x^3$\,.
Whereas \emph{adiabatic approximation}, \emph{singular perturbation}, and \emph{multiple scales} set $\dot y=0$ whence $y\approx x^2$ and thus predict only the linear growth of $\dot x= \epsilon x$\,.
Our normal form transforms get the deterministic dynamics correctly.
But what happens for stochastic dynamics?


Start by loading the procedure.
\begin{reduce}
in_tex "../stoNormForm.tex"$
\end{reduce}
Execute the construction of a normal form for this system.
Multiply a cubic terms in the $x$~\sde\ in order to count orders of approximation best (since the right-hand side is multiplied by~\verb|small|).
Multiply the bifurcation parameter by~\verb|small| in order to make it scale with~\(\eps^2\). 
\begin{reduce}
stonormalform(
    {small*epsilon*x(1)+small*x(1)^3
        -x(1)*y(1)*(1-small*w(1))},
    {-y(1)+x(1)^2+y(1)^2+y(1)*w(1)},
    {},
    5)$
%end;% optionally finish code here
\end{reduce}

With the above artifices, the procedure analyses the following system which reduce to the given one for \(\eps=1\)\,:
\begin{align*}&
\dot x_{1}=\sigma  \eps w_{1} x_{1} y_{1}+\eps^{2} \big(x_{1}^{3}+x_{1} 
\epsilon \big)-\eps x_{1} y_{1}
\\&
\dot y_{1}=\sigma  w_{1} y_{1}+\eps \big(x_{1}^{2}+y_{1}^{2}\big)-y_{1}
\end{align*}


\paragraph{Time dependent coordinate transform}
This transform is quite complicated, due to the noise, and involve fast-time future and history integrals. 
The deterministic terms at the end.
\begin{align*}&
y_{1}=\ParMath{\sigma  \eps^{3} \big(-4 \ou\big(\ou\big(w_{1},tt,1\big),tt,1\big)
 X_{1}^{2} Y_{1}^{2}+4 \ou\big(\ou\big(w_{1},tt,-1\big),tt,-1\big) X_{1}
^{4}-2 \ou\big(\ou\big(w_{1},tt,-1\big),tt,-1\big) X_{1}^{2} \epsilon +2
 \ou\big(w_{1},tt,2\big) X_{1}^{2} Y_{1}^{2}-10 \ou\big(w_{1},tt,1\big) 
X_{1}^{2} Y_{1}^{2}-3 \ou\big(w_{1},tt,1\big) Y_{1}^{4}+\ou\big(w_{1},tt
,-1\big) X_{1}^{4}+3 \ou\big(w_{1},tt,-1\big) X_{1}^{2} Y_{1}^{2}-2 \ou
\big(w_{1},tt,-1\big) X_{1}^{2} \epsilon \big)+\sigma  \eps^{2} \big(2 
\ou\big(w_{1},tt,1\big) Y_{1}^{3}-2 \ou\big(w_{1},tt,-1\big) X_{1}^{2} Y
_{1}\big)+\sigma  \eps \big(-\ou\big(w_{1},tt,1\big) Y_{1}^{2}+\ou\big(w
_{1},tt,-1\big) X_{1}^{2}\big)+\eps^{3} \big(X_{1}^{4}-7 X_{1}^{2} Y_{1}
^{2}-2 X_{1}^{2} \epsilon -Y_{1}^{4}\big)+\eps^{2} Y_{1}^{3}+\eps \big(X
_{1}^{2}-Y_{1}^{2}\big)+Y_{1}
}
\\&
x_{1}=\ParMath{\sigma  \eps^{3} \big(-\ou\big(w_{1},tt,3\big) X_{1} Y_{1}^{3}+\ou
\big(w_{1},tt,2\big) X_{1} Y_{1}^{3}+3 \ou\big(w_{1},tt,1\big) X_{1}^{3}
 Y_{1}\big)+\sigma  \eps^{2} \big(\ou\big(w_{1},tt,2\big) X_{1} Y_{1}^{2
}-\ou\big(w_{1},tt,1\big) X_{1} Y_{1}^{2}+\ou\big(w_{1},tt,-1\big) X_{1}
^{3}\big)+2 \eps^{3} X_{1}^{3} Y_{1}+\eps X_{1} Y_{1}+X_{1}
}
\end{align*}

\paragraph{Result normal form DEs}
\begin{align*}&
\dot Y_{1}=\ParMath{\sigma ^{2} \eps^{4} \big(8 \ou\big(\ou\big(w_{1},tt,-1\big),
tt,-1\big) w_{1} X_{1}^{4} Y_{1}-4 \ou\big(\ou\big(w_{1},tt,-1\big),tt,
-1\big) w_{1} X_{1}^{2} Y_{1} \epsilon +6 \ou\big(w_{1},tt,1\big) w_{1} 
X_{1}^{4} Y_{1}+22 \ou\big(w_{1},tt,-1\big) w_{1} X_{1}^{4} Y_{1}-4 \ou
\big(w_{1},tt,-1\big) w_{1} X_{1}^{2} Y_{1} \epsilon \big)+2 \sigma ^{2}
 \eps^{2} \ou\big(w_{1},tt,-1\big) w_{1} X_{1}^{2} Y_{1}+\sigma  \eps^{4
} \big(22 w_{1} X_{1}^{4} Y_{1}-4 w_{1} X_{1}^{2} Y_{1} \epsilon \big)+2
 \sigma  \eps^{2} w_{1} X_{1}^{2} Y_{1}+\sigma  w_{1} Y_{1}+\eps^{4} 
\big(6 X_{1}^{4} Y_{1}-4 X_{1}^{2} Y_{1} \epsilon \big)+4 \eps^{2} X_{1}
^{2} Y_{1}-Y_{1}
}
\\&
\dot X_{1}=\ParMath{-3 \sigma ^{2} \eps^{4} \ou\big(w_{1},tt,-1\big) w_{1} X_{1}
^{5}-2 \sigma  \eps^{4} w_{1} X_{1}^{5}+\eps^{4} \big(-X_{1}^{5}+2 X_{1}
^{3} \epsilon \big)+\eps^{2} X_{1} \epsilon 
}
\end{align*}
\begin{itemize}
\item As expected, \(Y_1=0\) is the stochastic slow manifold, and is exponentially attractive (almost always) in some domain about the origin.
\item The slow \(X_1\) evolution is independent of~\(Y_1\).  
Deterministically (\(\sigma=0\)), we predict a bifurcation to \(X_1\approx \pm\epsilon^{1/4}\).  The noise appears to modify this slightly.
\item The time-dependent coordinate transform maps these predictions back into the \(xy\)-plane.
\end{itemize}



\subsubsection{Focus on the slow manifold}

To directly construct the non-autonomous\slash stochastic slow centre manifold, just add the extra fifth parameter \verb|cman| when invoking the function.
\begin{reduce}
stonormalform(
    {small*epsilon*x(1)+small*x(1)^3
        -x(1)*y(1)*(1-small*w(1))},
    {-y(1)+x(1)^2+y(1)^2+y(1)*w(1)},
    {},
    5,cman)$
end;
\end{reduce}

\paragraph{Time dependent centre manifold coordinates}
This parametrisation is just the \(Y_1=0\) version of the full coordinate transform, but derived much more quickly.
\begin{align*}&
y_{1}=\ParMath{\sigma  \eps^{3} \big(4 \ou\big(\ou\big(w_{1},tt,-1\big),tt,-1
\big) X_{1}^{4}-2 \ou\big(\ou\big(w_{1},tt,-1\big),tt,-1\big) X_{1}^{2} 
\epsilon +\ou\big(w_{1},tt,-1\big) X_{1}^{4}-2 \ou\big(w_{1},tt,-1\big) 
X_{1}^{2} \epsilon \big)+\sigma  \eps \ou\big(w_{1},tt,-1\big) X_{1}^{2}
+\eps^{3} \big(X_{1}^{4}-2 X_{1}^{2} \epsilon \big)+\eps X_{1}^{2}
}
\\&
x_{1}=\ParMath{\sigma  \eps^{2} \ou\big(w_{1},tt,-1\big) X_{1}^{3}+X_{1}
}
\end{align*}

\paragraph{Result centre manifold DEs}
On the slow manifold the evolution is as in the full construction.
\begin{align*}&
\dot X_{1}=\ParMath{-3 \sigma ^{2} \eps^{4} \ou\big(w_{1},tt,-1\big) w_{1} X_{1}
^{5}-2 \sigma  \eps^{4} w_{1} X_{1}^{5}+\eps^{4} \big(-X_{1}^{5}+2 X_{1}
^{3} \epsilon \big)+\eps^{2} X_{1} \epsilon 
}
\end{align*}

